### 2020-03-19 abstract.tex
break apart into sections -- i think this makes it easier to parallelize

\begin{abstract}
    Abstracts are scary.
\end{abstract}