### 2020-03-19 intro.tex
Updated Intro; first BAD pass

\section{Introduction}

Imagine asking your computer to write a function for you and it works on the
first try.
%
Well, program synthesis isn't quite there.
%
But getting there will necessarily involve some back and forth between the user
and computer.
%
Often the function specification is less precise the than function requested.
%
In this case, a program synthesizer can come up with any number of solutions
that all meet the specification, to varying degrees of utility.

Our work is based on\textsc{Hoogle+}, a program synthesis engine for the
Haskell programming language \cite{hoogle_plus_2020}.
%
A user inputs a type signature of a function, like \texttt{a -> [Maybe a] -> a}
(read: ``A function that takes a thing and a list of optional-things, and
returns a thing'').
%
\textsc{Hoogle+} then looks for Haskell programs that use library calls (from
a small, builtin set of functions) that satisfies the type signature.
%
Note that there may be an infinite number of so-called \emph{inhabitants} of a type.
%
This presents a search problem for users of the system.

How can a user sift through the programs synthesized to find the right
implementation of the type signature, for their use case?
%
We extend the system to provide some input-output examples for a chosen set of
queries, enough to run our study.
%
For now these examples were stubbed and chosen without particular regard for how
they might distinguish various candidate solutions.
%