### 2020-03-19 discussion.tex
Task 2 light discussion; future work created

\section{Discussion}
\label{sec:discussion}

\subsection{Statistical insignificance}
Why it be like it do?

\subsection{Task 2}
The Wilcoxon Rank-Sum test on task 2 was $p=0.03$, showing a runaway success
that examples do help participants find the correct program faster.
%
We suspect that second task was unique among the four participants completed.
%
The tool produced up to ten candidate programs for a given type signature, and
for task 2, the correct solution was at the very end.
%
Task 1 and 4 had the correct solution listed in the first and second places,
respectively.