### 2020-03-19 conclusion.tex
added conclusion

\section{Conclusion}

We have shown an experimental setup to suss out when input-output examples help
users in a type-directed synthesis environment.
%
Overall our results did not show a statistical significance ($p = 0.14$);
however, we did find some interesting data points for further research.
%
Specifically, one task did conclude a statistically significant result ($p =
0.03$), Task 2.
%
We offered a theory explaining it and future experiments that could falsify our
claim.
%
We found that users did rely on an interpreter more in the absence of examples
to offer a "feel" for the candidate programs.