### 2020-03-19 abstract.tex
abstract and methods update

\begin{abstract}
Program synthesis has been focusing on making newer and better tools; however,
few have evaluated how people like to interact with them and what features users
find useful.
%
We evaluate the addition of examples to type directed program synthesis.
%
We measured 10 participants between two variants, each with 4 tasks.
%
We do not find a statistical significance, in general, with the addition of
examples.
%
We do note, that one task did have a significant difference in performance
between groups.
%
Further evaluation is needed to determine a reason for this.
\end{abstract}