### 2020-03-19 future.tex
whoops duplicates

\section{Future Work}
\label{sec:future}

This work touches on but does not seek to evaluate whether recognition of a
correct solution or recall of necessary parts of a correct solution are more
useful in this type-directed search setting.
%
In our \noexamples group, participants largely relied on the interpreter to tell
them what different parts of a candidate solution were doing.
%
In effect, they used the interpreter as a source of documentation.
%
So, it remains to be seen what kinds of tasks users accomplish faster with
documentation as an aid vs input-output examples.
%