### 2020-03-20 intro.tex

\section{Introduction}

Program Synthesis works by taking a specification and generating a program that
implements the specification.
%
In a sense it is the holy grail of computer science.
%
There is a plethora of work trying to go straight to the goal.
%
Some use input-output examples as their specification
\cite{Feser_Chaudhuri_Dillig_2015}.
%
Others use types and examples \cite{Osera_Zdancewic_2015}.
%
Still others use just types \cite{hoogle_plus_2020}.

However, in lieu of waiting for some graduate student to solve this field, we
adopt a more practical approach and want to bring existing techniques to users.
%
In doing so, we seek to help users understand the solutions that program
synthesizers come up with.
%
Since current specifications tend to be underpowered, many unhelpful programs
come up from synthesizers.
%

Specifically, we seek to understand the degree to which input-output examples,
as properties of synthesized functions (instead of as specifications), aid users
in finding the program they came to the synthesizer for.

We base our work on \hoogleplus, a program synthesis engine for the
Haskell programming language \cite{hoogle_plus_2020}.
%
A user inputs a type signature of a function, like \texttt{a -> [Maybe a] -> a}
(read: ``A function that takes a thing and a list of optional-things, and
returns a thing'').
%
\hoogleplus then looks for Haskell programs that use library calls (from
a small, builtin set of functions) that satisfies the type signature.
%
Note that there may be an infinite number of candidate solutions that satisfy a type.
%
This presents a search problem for users of the system.

How can a user sift through the programs synthesized to find the right
implementation of the type signature, for their use case?
%
We extend the system to provide some input-output examples for a chosen set of
queries, enough to run our study.
%
For now these examples were stubbed and chosen without particular regard for how
they might distinguish various candidate solutions.
%