### 2020-03-20 conclusion.tex
Really just remove the word suss. It's gross

\section{Conclusion}

We have demonstrated an experimental setup to evaluate whether input-output
examples help users in a type-directed synthesis environment.
%
Overall our results did not show a statistical significance ($p = 0.14$);
however, we did find some interesting data points for further research.
%
Specifically, one task did conclude a statistically significant result ($p =
0.03$), Task~2.
%
We offered a theory explaining it and future experiments that could falsify our
claim.
%
We found that users did rely on an interpreter more in the absence of examples
to offer a "feel" for the candidate programs and to lookup documentation.
%
We believe our study was sufficiently underpowered to be unable to make a
definitive claim on the overall utility of examples in a type-directed synthesis
environment.